---
title: Graphical Testing 
output:
  html_document:
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
library(gsDesign)
library(gMCPLite)
library(dplyr)
library(kableExtra)
```

## Multiplicity diagram for hypothesis testing

```{r}
nameHypotheses <- c(<%=paste0("\"", paste0(input$hypothesesMatrix[,"Name"], collapse = "\", \""), "\"")%>)
alphaHypotheses <- c(<%=paste0(sapply(input$hypothesesMatrix[,"Alpha"], arithmetic_to_numeric), collapse = ", ")%>)
FWER <- sum(alphaHypotheses)
GroupHypotheses <- c(<%=paste0("\"", paste0(unique(input$hypothesesMatrix[,"Group"]), collapse = "\", \""), "\"")%>)
n_hypo <- <%=nrow(input$hypothesesMatrix)%>
m <- matrix(c(<%=paste0(c(df2graph(namesH = input$hypothesesMatrix[,1], df = data.frame(input$trwtMatrix)[(data.frame(input$trwtMatrix)[,1] %in% input$hypothesesMatrix[,1]) & (data.frame(input$trwtMatrix)[,2] %in% input$hypothesesMatrix[,1]),])), collapse = ", ")%>), nrow = n_hypo)

h <- hGraph(
  nHypotheses     = n_hypo,
  nameHypotheses  = nameHypotheses,
  alphaHypotheses = alphaHypotheses ,
  m               = m,
  fill            = factor(GroupHypotheses, levels = unique(GroupHypotheses)),
  palette         = c(<%=paste0("\"", paste0(hgraph_palette(pal_name = rv_nodes$pal_name, n = length(unique(input$hypothesesMatrix[,"Group"])), alpha = rv_nodes$pal_alpha), collapse = "\", \""), "\"")%>),
  labels          = unique(GroupHypotheses),
  legend.name     = <%=paste0("\"", input$legend.name, "\"")%>,
  legend.position = <%=paste0("\"", input$legendPosition, "\"")%>,
  halfWid         = <%=rv_nodes$width%>,
  halfHgt         = <%=rv_nodes$height%>,
  trhw            = <%=rv_edges$trhw%>,
  trhh            = <%=rv_edges$trhh%>,
  trprop          = <%=rv_edges$trprop%>,
  digits          = <%=rv_nodes$digits%>,
  trdigits        = <%=rv_edges$trdigits%>,
  size            = <%=rv_nodes$size%>,
  boxtextsize     = <%=rv_edges$boxtextsize%>,
  legend.textsize = <%=input$legend.textsize%>,
  arrowsize       = <%=rv_edges$arrowsize%>,
  offset          = <%=rv_edges$offset%>,
  x               = <%=if(is.null(input$nodeposMatrix[,"x"]) | !setequal(input$nodeposMatrix[,"Hypothesis"], input$hypothesesMatrix[,"Name"])){"NULL"} else{paste0("c(", paste0(as.numeric(input$nodeposMatrix[,"x"]), collapse = ", "), ")")}%>,
  y               = <%=if(is.null(input$nodeposMatrix[,"y"]) | !setequal(input$nodeposMatrix[,"Hypothesis"], input$hypothesesMatrix[,"Name"])){"NULL"} else{paste0("c(", paste0(as.numeric(input$nodeposMatrix[,"y"]), collapse = ", "), ")")}%>,
  wchar           = <%=paste0("\"", rv_nodes$wchar, "\"")%>
)<%= if (input$plotTitle != "") { paste0(" +\n  labs(title = \"", input$plotTitle, "\") +", "\n  theme(plot.title = element_text(size = ", input$title.textsize, ", hjust = ", input$title.position, "))") } %><%= if (input$plotCaption != "") " +" -%>
<%= if (input$plotCaption != "") { paste0("\n  labs(caption = \"", input$plotCaption, "\") +", "\n  theme(plot.caption = element_text(size = ", input$caption.textsize, ", hjust = ", input$caption.position, "))") } %>

plot(h)
```

## Original design for each hypothesis{.tabset}

```{r, results='asis'}
endpoint<-c(<%=paste0("\"",paste0(sapply(1:nrow(input$hypothesesMatrix),function(i){ifelse(input[[paste0("design_type_",i)]]=='fix','fix','gsd')}),collapse="\", \""), "\"")%>)
k <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$k}),collapse=", ")%>)
test.type <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$test.type}),collapse=", ")%>)
alpha <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$alpha}),collapse=", ")%>)
beta <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$beta}),collapse=", ")%>)
astar <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$astar}),collapse=", ")%>)
timing <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$timing}),collapse=", ")%>)
sfu <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$upper$sf}),collapse=", ")%>)
sfupar <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$upper$param}),collapse=", ")%>)
sfl <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$lower$sf}),collapse=", ")%>)
sflpar <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$lower$param}),collapse=", ")%>)
lambdaC <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$lambdaC}),collapse=", ")%>)
hr <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$hr}),collapse=", ")%>)
hr0 <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$hr0}),collapse=", ")%>)
etaC <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$etaC}),collapse=", ")%>)
etaE <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$etaE}),collapse=", ")%>)
gamma <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$gamma}),collapse=", ")%>)
R <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$R}),collapse=", ")%>)
S <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$S}),collapse=", ")%>)
T <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{max(GetDesign()[[i]]$T)}),collapse=", ")%>)
minfup <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$minfup}),collapse=", ")%>)
ratio <- list(<%=paste0(sapply(1:nrow(input$hypothesesMatrix),function(i) if (input[[paste0("design_type_",i)]]=='fix'){NULL}else{GetDesign()[[i]]$ratio}),collapse=", ")%>)

gsDlist<-list()
gsDlist <- sapply(1:n_hypo, function(i) if (endpoint[i]=='fix'){NULL}else {
  gsDesign::gsSurv(
        k = k[[i]],
        test.type = test.type[[i]],
        alpha = alpha[[i]],
        beta = beta[[i]],
        astar = astar[[i]],
        timing = timing[[i]],
        sfu = sfu[[i]],
        sfupar = sfupar[[i]],
        sfl = sfl[[i]],
        sflpar = sflpar[[i]],
        lambdaC = lambdaC[[i]],
        hr = hr[[i]],
        hr0 = hr0[[i]],
        eta = etaC[[i]],
        etaE = etaE[[i]],
        gamma = gamma[[i]],
        R = R[[i]],
        S = S[[i]],
        T = T[[i]],
        minfup = minfup[[i]],
        ratio = ratio[[i]])
})
for (i in 1:n_hypo) {
  cat("###", paste(nameHypotheses[i], " \n"))
  d <- gsDlist[[i]]
  # If not group sequential for this hypothesis, print the max alpha allocated
  # and the nominal p-value
  if (is.null(d)) {
    cat(paste0(nameHypotheses[i], " is fixed sample size design."))
    cat("\n\n")
  }
  # For group sequential tests, print max alpha allocated and
  # corresponding group sequential bounds
  if (!is.null(gsDlist[[i]])) {
    tab <- gsDesign::gsBoundSummary(gsDlist[[i]])
    rownames(tab) <- 1:nrow(tab)
    kable(tab, caption = paste0("Design for ", nameHypotheses[i]), row.names = FALSE) %>% kable_styling() %>% cat()
    cat("\n\n")
    }
}
```

## Result entries at the timing of analysis

```{r}
Name<-c(<%=paste0("\"",paste0(sapply(sapply(seq(nrow(input$hypothesesMatrix)),function(i) if (input[[paste0("design_type_", i)]]=='fix'){input$hypothesesMatrix[i,'Name']}else{rep(input$hypothesesMatrix[i,'Name'],nrow(input[[paste0("pvalMatrix_",i)]]))}),paste,collapse='\","'),collapse = "\", \""), "\"")%>)
Group<-c(<%=paste0("\"",paste0(sapply(sapply(seq(nrow(input$hypothesesMatrix)),function(i) if (input[[paste0("design_type_", i)]]=='fix'){input$hypothesesMatrix[i,'Group']}else{rep(input$hypothesesMatrix[i,'Group'],nrow(input[[paste0("pvalMatrix_",i)]]))}),paste,collapse='\","'),collapse = "\", \""), "\"")%>)
ObsPval<-c(<%=paste0(sapply(sapply(seq(nrow(input$hypothesesMatrix)),function(i) if (input[[paste0("design_type_", i)]]=='fix'){input[[paste0("pval_", i)]]}else{as.numeric(input[[paste0("pvalMatrix_", i)]][,3])}),paste,collapse=", "),collapse=", ")%>)
ObsEvnts<-c(<%=paste0(sapply(sapply(seq(nrow(input$hypothesesMatrix)),function(i) if (input[[paste0("design_type_", i)]]=='fix'){NA}else{as.numeric(input[[paste0("pvalMatrix_", i)]][,2])}),paste,collapse=", "),collapse=", ")%>)
Analysis<-c(<%=paste0(sapply(sapply(seq(nrow(input$hypothesesMatrix)),function(i) if (input[[paste0("design_type_", i)]]=='fix'){NA}else{as.numeric(input[[paste0("pvalMatrix_", i)]][,1])}),paste,collapse=", "),collapse=", ")%>)

n.Iplan<- c(<%=paste0(sapply(sapply(seq(nrow(input$hypothesesMatrix)),function(i) if (input[[paste0("design_type_", i)]]=='fix'){NA}else{GetDesign()[[i]]$n.I[1:nrow(input[[paste0("pvalMatrix_",i)]])]}),paste,collapse=", "),collapse=", ")%>)
n.Imax <- c(<%=paste0(sapply(sapply(seq(nrow(input$hypothesesMatrix)),function(i) if (input[[paste0("design_type_", i)]]=='fix'){NA}else{rep(max(GetDesign()[[i]]$n.I),nrow(input[[paste0("pvalMatrix_",i)]]))}),paste,collapse=", "),collapse=", ")%>)

spending_method <- c(<%=paste0("\"",paste0(sapply(sapply(seq(nrow(input$hypothesesMatrix)),function(i) if (input[[paste0("design_type_", i)]]=='fix'){"fix"}else{rep(input[[paste0("spending_interim_", i)]],nrow(input[[paste0("pvalMatrix_",i)]]))}),paste,collapse='\","'),collapse = "\", \""), "\"")%>)

spendingTime<-ifelse(spending_method=='fix',NA,ifelse(spending_method=='info',ObsEvnts/n.Imax,pmin(ObsEvnts,n.Iplan)/n.Imax))
inputResults<-data.frame(Name,Group,Analysis,n.Iplan,ObsEvnts,ObsPval,spendingTime)
kable(inputResults, caption = "Results for Interim Analysis.") %>%
  kable_styling()
```

## Testing hypotheses
### Compute sequential p-values for each hypothesis
```{r}
EOCtab <- inputResults %>%
  group_by(Name) %>%
  slice(1) %>%
  ungroup() %>%
  select("Name","Group","ObsPval")
EOCtab$seqp <- .9999
for (i in 1:n_hypo) {
  EOCtab$seqp[i] <-
    ifelse(is.null(gsDlist[[i]]), EOCtab$ObsPval[i], {
      tem <- filter(inputResults, Name == unique(inputResults$Name)[i])
      sequentialPValue(
        gsD = gsDlist[[i]], interval = c(.0001, .9999),
        n.I = tem$ObsEvnts,
        Z = -qnorm(tem$ObsPval),
        usTime = tem$spendingTime
      )
    })
}
EOCtab <- EOCtab %>% select(-"ObsPval")
```

### Evaluate hypothesis rejection using gMCPLite
```{r,message=FALSE,warning=FALSE}
# Make a graph object
graph <- matrix2graph(m)
# Add weights to the object based on alpha allocation
graph <- gMCPLite::setWeights(gMCPLite::matrix2graph(m), alphaHypotheses / FWER)
result <- gMCP(graph = graph, pvalues = EOCtab$seqp, alpha = FWER)
# now map back into EOCtable
EOCtab$Rejected <- result@rejected
EOCtab$adjPValues <- result@adjPValues
```

## Verification of hypotheses rejected

```{r}
# Number of graphs is used repeatedly
ngraphs <- length(result@graphs)
# Set up tibble with hypotheses rejected at each stage
rejected <- NULL
for (i in 1:length(result@graphs)) {
  rejected <- rbind(
    rejected,
    tibble(
      Name = 1:n_hypo, Stage = i,
      Rejected = as.logical(result@graphs[[i]]@nodeAttr$rejected)
    )
  )
}
rejected <- rejected %>%
  filter(Rejected) %>%
  group_by(Name) %>%
  summarize(graphRejecting = min(Stage) - 1, .groups = "drop") %>% # Last graph with weight>0 where H rejected
  arrange(graphRejecting)
# Get final weights
# for hypotheses not rejected, this will be final weight where
# no hypothesis could be rejected
lastWeights <- as.numeric(result@graphs[[ngraphs]]@weights)
lastGraph <- rep(ngraphs, nrow(EOCtab))
# We will update for rejected hypotheses with last positive weight for each
if (ngraphs > 1) {
  for (i in 1:(ngraphs - 1)) {
    lastWeights[rejected$Name[i]] <- as.numeric(result@graphs[[i]]@weights[rejected$Name[i]])
    lastGraph[rejected$Name[i]] <- i
  }
}
EOCtab$lastAlpha <- FWER * lastWeights
EOCtab$lastGraph <- lastGraph
EOCtabx <- EOCtab
names(EOCtabx) <- c(
  "Name", "Group", "Sequential p",
  "Rejected", "Adjusted p", "Max alpha allocated", "Last Graph"
)
# Display table with desired column order
# Delayed following until after multiplicity graph sequence
# EOCtabx %>% select(c(1:4,7,5:6,8)) %>% kable() %>% kable_styling()
```

### Multiplicity graph sequence from gMCPLite {.tabset}

```{r,message=FALSE,warning=FALSE,results='asis'}
for (i in 1:ngraphs) {
  mx <- result@graphs[[i]]@m
  rownames(mx) <- NULL
  colnames(mx) <- NULL
  g <- gMCPLite::hGraph(
    nHypotheses = n_hypo,
    alphaHypotheses = result@graphs[[i]]@weights * FWER,
    m = mx,
    nameHypotheses = nameHypotheses,
    palette         = c(<%=paste0("\"", paste0(hgraph_palette(pal_name = rv_nodes$pal_name, n = length(unique(input$hypothesesMatrix[,"Group"])), alpha = rv_nodes$pal_alpha), collapse = "\", \""), "\"")%>),
    labels          = unique(GroupHypotheses),
    legend.name     = <%=paste0("\"", input$legend.name, "\"")%>,
    legend.position = <%=paste0("\"", input$legendPosition, "\"")%>,
    halfWid         = <%=rv_nodes$width%>,
    halfHgt         = <%=rv_nodes$height%>,
    trhw            = <%=rv_edges$trhw%>,
    trhh            = <%=rv_edges$trhh%>,
    trprop          = <%=rv_edges$trprop%>,
    digits          = <%=rv_nodes$digits%>,
    trdigits        = <%=rv_edges$trdigits%>,
    size            = <%=rv_nodes$size%>,
    boxtextsize     = <%=rv_edges$boxtextsize%>,
    legend.textsize = <%=input$legend.textsize%>,
    arrowsize       = <%=rv_edges$arrowsize%>,
    offset          = <%=rv_edges$offset%>,
    x               = <%=if(is.null(input$nodeposMatrix[,"x"]) | !setequal(input$nodeposMatrix[,"Hypothesis"], input$hypothesesMatrix[,"Name"])){"NULL"} else{paste0("c(", paste0(as.numeric(input$nodeposMatrix[,"x"]), collapse = ", "), ")")}%>,
    y               = <%=if(is.null(input$nodeposMatrix[,"y"]) | !setequal(input$nodeposMatrix[,"Hypothesis"], input$hypothesesMatrix[,"Name"])){"NULL"} else{paste0("c(", paste0(as.numeric(input$nodeposMatrix[,"y"]), collapse = ", "), ")")}%>,
    wchar           = <%=paste0("\"", rv_nodes$wchar, "\"")%>
) 
  cat(" \n")
  cat("####", paste(" Graph", as.character(i), " \n\n"))
  print(g)
  cat(" \n\n\n")
}
```

### Comparison of sequential p-values to multiplicity graphs

We can compare sequential p-values to available $\alpha$ in each graph.
In the column 'Last Graph' we can see one of 2 things:

1) For rejected hypotheses, the maximum $\alpha$ allocated to the hypothesis. 
2) The maximum $\alpha$ allocated to the non-rejected hypotheses; since each sequential p-value is greater than the allocated $\alpha$ for the corresponding hypothesis, none of these hypotheses were rejected.

```{r}
EOCtabx %>%
  select(c(1:3, 6, 4:5, 7)) %>%
  kable() %>%
  kable_styling()
```

### Bounds at final $\alpha$ allocated for group sequential tests {.tabset}

As a separate validation, we examine group sequential bounds for each hypothesis updated with 1) the maximum $\alpha$ allocated above, 2) the number of events at each analysis, and 3) the cumulative spending at each analysis above. The nominal p-value for at least one of the analyses performed for each rejected hypotheses should be less than or equal to the nominal p-value in the group sequential design.
For each hypothesis not rejected, all nominal p-values are greater than the its corresponding bound.
For hypotheses tested without a group sequential design, the nominal p-value for the test of that hypothesis can be compared to the maximum alpha allocated in the above table.

```{r,results='asis'}
for (i in 1:n_hypo) {
  cat("####", paste(nameHypotheses[i], " \n"))
  hresults <- inputResults %>% filter(Name == unique(inputResults$Name)[i])
  xx <- paste("Max alpha allocated from above table: ",
    as.character(EOCtab$lastAlpha[i]),
    sep = ""
  )
  d <- gsDlist[[i]]
  # If not group sequential for this hypothesis, print the max alpha allocated
  # and the nominal p-value
  if (is.null(d)) {
    cat("Maximum alpha allocated: ")
    cat(EOCtab$lastAlpha[i])
    cat("\n\n")
    cat("Nominal p-value for hypothesis test: ")
    cat(hresults$ObsPval)
    cat("\n\n")
  }
  # For group sequential tests, print max alpha allocated and
  # corresponding group sequential bounds
  if (!is.null(gsDlist[[i]])) {
    cat("Nominal p-values at each analysis for comparison to bounds in table below:\n\n")
    cat(hresults$ObsPval)
    cat("\n\n")
    # Get other info for current hypothesis
    events<-hresults$ObsEvnts
    length(events)<-d$k
    length(spendingTime)<-d$k
    n.I <- ifelse(is.na(events),d$n.I,events)
    usTime <- ifelse(is.na(spendingTime),d$timing,spendingTime)
    n.Imax <- max(d$n.I)
    # If no alpha allocated, just print text line to note this along with the 0 alpha allocated
    if (EOCtab$lastAlpha[i] == 0) {
      cat("Maximum alpha allocated: 0\n\n")
      cat("No testing required\n\n")
    }
    if (EOCtab$lastAlpha[i] > 0) {
      dupdate <- gsDesign::gsDesign(
        alpha = EOCtab$lastAlpha[i],
        k = d$k,
        n.I = n.I,
        usTime = usTime,
        maxn.IPlan = n.Imax,
        n.fix = d$n.fix,
        test.type = 1,
        sfu = d$upper$sf,
        sfupar = d$upper$param
      )
      tabl <- gsDesign::gsBoundSummary(dupdate,
        Nname = "Events",
        exclude = c(
          "B-value", "CP", "CP H1", "Spending",
          "~delta at bound", "P(Cross) if delta=0",
          "PP", "P(Cross) if delta=1"
        )
      )
      kable(tabl, caption = xx, row.names = FALSE) %>%
        kable_styling() %>%
        cat()
      cat("\n\n")
    }
  }
}
```

## Environment information

-   Run date: `r Sys.time()`
-   R version: `r R.version.string`
